---
title: "Data cleaning for cognition ExWAS and SVI interaction analyses"
author: "Erika Walker"
date: "`r Sys.Date()`"
output: html_document
---

Purpose: create a dataset of NHANES cycles 1, 7, and 8 (1999-2000, 2011-2012, and 2013-2014) containing the variables needed for the cognition ExWAS and cognition SVI EWIS analyses; clean the chemical exposure dataset

Participant inclusion criteria: in cycles 1, 7, or 8 (cycles with both cognitive measures and fish consumption measures), aged 60+, with at least one chemical exposure measurement

Data sources: 

https://www.kaggle.com/datasets/nguyenvy/nhanes-19882018 - will need to download the full RDATA environment, plus the following files individually: dictionary_nhanes, chemicals_clean, comments_clean, weights_clean

https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx - for additional variables not included in the above processed dataset (files are read in directly using haven::read_xpt() and do not need to be downloaded separately)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(sjlabelled)
```

# Load and prepare NHANES files
```{r}
# load source environment: note that this takes ~ 5GB memory
load('Data/w - nhanes_1988_2018.RData')

# remove files not needed
rm(list = setdiff(ls(), c('chemicals_clean', 'comments_clean', 'demographics_clean', 'occupation_clean', 
                          'questionnaire_clean', 'response_clean', 'weights_clean')))

# select variables needed from demographics, occupation, questionnaire, and response files
demographics_clean <- demographics_clean %>%
  select(SEQN, SEQN_new, SDDSRVYR, RIDAGEYR, RIAGENDR, RIDRETH1, INDFMPIR, DMDEDUC2, DMDEDUC3,
         SDMVPSU, SDMVSTRA, WTMEC2YR, WTMEC4YR, WTINT2YR, WTINT4YR) # identifiers, demographics, survey design

occupation_clean <- occupation_clean %>%
  select(SEQN, SEQN_new, SDDSRVYR, OCD150, OCQ380) # occupational status

questionnaire_clean <- questionnaire_clean %>%
  select(SEQN, SEQN_new, SDDSRVYR, ALQ101, ALQ120Q, ALQ120U) # alcohol consumption
  
response_clean <- response_clean %>%
  select(SEQN, SEQN_new, SDDSRVYR, URXUCR, BMXBMI, BMXWAIST, VNEGFR) # creatinine, BMI, waist circumference, eGFR

# merge the files together
nhanes_data <- left_join(demographics_clean,
                         occupation_clean,
                         by = c('SEQN', 'SEQN_new', 'SDDSRVYR')) %>%
    left_join(.,
            questionnaire_clean,
            by = c('SEQN', 'SEQN_new', 'SDDSRVYR')) %>%
  left_join(.,
            response_clean,
            by = c('SEQN', 'SEQN_new', 'SDDSRVYR'))
dim(nhanes_data) # 135310     24

# filter to cycles 1, 7, 8 (years 1999-2000, 2011-2012, 2013-2014)
nhanes_data <- nhanes_data %>%
  filter(SDDSRVYR %in% c(1, 7, 8))
dim(nhanes_data) # 29896    24

# create an initial indicator for adults 60+ - will provide a starting point for chemical cleaning
nhanes_data <- nhanes_data %>%
  mutate(aged_60plus = case_when(RIDAGEYR >= 60 ~ 1,
                                 .default = 0))

table(nhanes_data$aged_60plus, useNA = 'always')
# N = 5466

# pull string of SEQNs for these participants
aged_60plus <- nhanes_data %>%
  filter(aged_60plus == 1) %>%
  pull(SEQN)

# load data dictionary for chemicals: will match chemical variables to comment variables
chem_dictionary <- read_csv('Data/dictionary_nhanes.csv') %>%
  filter(in_dataset == 'Chemicals') %>%
  select(variable_codename_use, variable_description_use, unique_cycles,
         comment_codename_use, chemical_family)

# why 598 variables in chemicals_clean but 599 entries in dictionary?
#chem_find_extra <- chem_dictionary %>%
#  filter(!variable_codename_use %in% colnames(chemicals_clean))
# nothing missing - checked dictionary manually and there is a duplicate entry for LBXTC2
# remove duplicate
chem_dictionary <- chem_dictionary %>%
  distinct()

```

# Chemical and participant inclusion/exclusion

Chemical exclusion criteria: 

 - start with chemicals measured in the relevant cycles (1999-2000, 2011-2012, 2013-2014) --> 395
 
 - drop some chemical families based on Lauren's work --> 366
 
 - drop lipid-unadjusted variables when an adjusted version exists --> 316
 
 - drop chemicals if over 50% of measurements are below lower limit of detection --> 147
 
## Filter chemicals dataset to adults 60+ in relevant cycles
```{r}
chems <- chemicals_clean %>%
  filter(SDDSRVYR %in% c(1, 7, 8)) %>%
  filter(SEQN %in% aged_60plus)
dim(chems) # 5078  598

# note that aged_60plus is 5466 participants while chems contains 5078: not all participants have chemical measures so this is fine

```

## Preliminary exclusion: drop any chems not measured in the relevant cycles
```{r}

# identify chems measured in cycles 1, 7, 8
chem_dictionary_3cycles <- chem_dictionary %>%
  filter(grepl(paste(c('1999-2000', '2011-2012', '2013-2014'), collapse = '|'), unique_cycles))
dim(chem_dictionary_3cycles) # 398   5 --> 3 rows are for identifiers, so 395 chems were measured in the cycles being used

# reduce chems dataset to these variables
chems <- chems %>%
  select(chem_dictionary_3cycles$variable_codename_use)
dim(chems) # 5078  398 --> 3 identifiers, 395 chems

```

## Per Lauren's process: drop dietary components, phytoestrogens, and nicotine metabolites measured only in smokers
```{r}

nicotine_to_remove <- c('URXNOXT', 'URXCOXT', 'URXNNCT', 'URXANTT', 'URXANBT', 'URXNICT')
# see https://wwwn.cdc.gov/Nchs/Nhanes/2013-2014/UCOT_H.htm: 
# these chems were only measured if urinary cotinine > 20 ng/mL (as indicator of smoking)

dictionary_included <- chem_dictionary_3cycles %>%
  filter(!chemical_family %in% c('Dietary Components',
                                 'Phytoestrogens')) %>%
  filter(!variable_codename_use %in% nicotine_to_remove)
dim(dictionary_included) # 369   5 --> 3 identifiers, 366 chems

chems <- chems %>%
  select(dictionary_included$variable_codename_use)
dim(chems) # 5078  369 --> 366 chems

```

## Drop lipid-unadjusted chems
```{r}
# some vars have 2 versions: unadjusted and lipid-adjusted versions - want to keep the adjusted ones only
# use data dictionary to first identify the lipid-adjusted chems using the descriptions (variable_description_use)
# then identify the unadjusted versions to drop

# pull info for chems that have 'Lipid Adj' in the description
lipid_adj_dictionary <- dictionary_included %>%
  filter(grepl('Lipid Adj', variable_description_use))

# pull out just the chem names
lipid_adj_codenames <- data.frame(lipid_adj_dictionary$variable_codename_use)

# make a separate variable of the chem names, dropping the 'LA' to get the unadjusted versions
# example: LBX028LA -> LBX028
lipid_unadj_codenames <- lipid_adj_codenames %>%
  rename(variable_codename_use = lipid_adj_dictionary.variable_codename_use) %>%
  mutate(chem_codenames_unadj = gsub('LA$', '', variable_codename_use)) %>% # $ means search end of string
  select(chem_codenames_unadj)

# update dictionary and chems dataset to remove the lipid unadj variables
dictionary_included <- dictionary_included %>%
  filter(!variable_codename_use %in% lipid_unadj_codenames$chem_codenames_unadj)
dim(dictionary_included) # 319   5 --> 3 identifiers, 316 chems

chems <- chems %>%
  select(dictionary_included$variable_codename_use)
dim(chems) # 5078  319 --> 316 chems

```

## Drop chems with over 50% of measurements below limit of detection (LOD)

The chemical comments dataset indicates if measurements are above or below the chemical's LOD

If comment variable = 0, that measurement is above the LOD

If comment variable = 1, that measurement is below the LOD

```{r}
# filter comments dataset to participants aged 60+ in relevant cycles
chem_comments <- comments_clean %>% 
  filter(SDDSRVYR %in% c(1, 7, 8)) %>%
  filter(SEQN %in% aged_60plus)
dim(chem_comments) # 5078  505

# keep only the comment vars for currently included chems
comment_names <- dictionary_included %>%
  drop_na(comment_codename_use) %>%
  pull(comment_codename_use)

chem_comments <- chem_comments %>%
  select(SEQN, SEQN_new, SDDSRVYR, all_of(comment_names))
dim(chem_comments) # 5078  319 --> 3 identifiers, 316 comment variables

# process for determining % above LOD
# sum of count above (comment = 0)
# sum of count below (comment = 1)
# above + below = total participants with measurement
# above / total * 100 = % above LOD

# turn comments into a long dataset: one chem measurement/comment per row
long_comments <- chem_comments %>%
  pivot_longer(cols = all_of(comment_names), # specify all columns except the identifiers
               values_to = 'comment',
               names_to = 'comment_codename') %>%
  drop_na(comment) %>% # if the comment is NA, then that measurement doesn't exist for that participant - drop
  select(comment_codename, comment)
dim(long_comments) # 285518      2

# calculate number of measurements above vs below LOD for each chem
lod_calc <- long_comments %>%
  group_by(comment_codename) %>%
  summarize(n_above_lod = sum(comment == 0), 
            n_below_lod = sum(comment == 1),
            n_measures = n_above_lod + n_below_lod,
            pct_above_lod = (n_above_lod / n_measures) * 100) %>%
  mutate(use_chem = ifelse(pct_above_lod >= 50, 1, 0))

table(lod_calc$use_chem, useNA = 'always')

# look at just the chems to keep
above_50_pct <- lod_calc %>%
  filter(use_chem == 1)

# use the dictionary to match which comment codename goes to which chemical and filter it to just the chems to keep
dictionary_included <- dictionary_included %>%
  filter(comment_codename_use %in% above_50_pct$comment_codename)
dim(dictionary_included) # 147   5
# identifiers dropped off of dictionary here, so 147 chems now

# keep only these vars in the chems dataset
chems <- chems %>%
  select(SEQN, SEQN_new, SDDSRVYR, all_of(dictionary_included$variable_codename_use))
dim(chems) # 5078  150 --> 3 identifiers, 147 chems

```

## Drop participants without any chem measures
```{r}
# starting N = 5078 (chems dataset)
# create long dataset, one row per measure, in order to drop off all NAs together
drop_chem_nas <- chems %>%
  pivot_longer(cols = all_of(dictionary_included$variable_codename_use),
               names_to = 'chem',
               values_to = 'measure') %>%
  drop_na(measure)
dim(drop_chem_nas) # 147790      5

# pull sample size of participants with at least one measure
print(length(unique(drop_chem_nas$SEQN)))
# N = 5078 -> N = 4982

# keep only these participants in chems dataset
chems <- chems %>%
  filter(SEQN %in% unique(drop_chem_nas$SEQN))
dim(chems) # 4982  150

```

## Update analysis inclusion variable based on chemical measure availability
```{r}
# current nhanes_data$aged_60plus N = 5466
# chems participant N = 4982 (additionally dropped anyone without chem data)

included_seqns <- chems$SEQN

nhanes_data <- nhanes_data %>%
  mutate(include = case_when(SEQN %in% included_seqns ~ 1,
                             .default = 0))

table(nhanes_data$include, useNA = 'always')
table(nhanes_data$aged_60plus, nhanes_data$include, useNA = 'always')
# confirmed N = 4982 to include

```

## Add serum cotinine to nhanes_sample
```{r}
# pull from chemicals_clean to get measures for full sample, not just included
cotinine <- chemicals_clean %>%
  filter(SDDSRVYR %in% c(1, 7, 8)) %>% 
  select(SEQN, SDDSRVYR, LBXCOT)

nhanes_data <- left_join(nhanes_data,
                         cotinine,
                         by = c('SEQN', 'SDDSRVYR'))

summary(nhanes_data$LBXCOT)

```

# Save the included chems dictionary
The dictionary will be used later on to identify the chems to use when running models
```{r}

# add LOD % info of each chem based on included sample to dictionary
dictionary_included <- left_join(dictionary_included,
                                 above_50_pct,
                                 join_by(comment_codename_use == comment_codename))

# reorganize
dictionary_included <- dictionary_included %>%
  select(-use_chem) %>%
  relocate(n_measures, .before = n_above_lod) %>%
  relocate(unique_cycles, .after = pct_above_lod) %>%
  relocate(comment_codename_use, .after = unique_cycles)

# save dictionary
write_csv(dictionary_included, file = 'Data/use_these_chems.csv')

```

# Import cognitive data (outcome)

Using Digit Symbol Substitution Test (DSST) score
```{r}
# Load files from NHANES site
cfq1 <- read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/1999-2000/CFQ.XPT')
cfq7 <- read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/2011-2012/CFQ_G.XPT')
cfq8 <- read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/2013-2014/CFQ_H.XPT')

cfq_all <- bind_rows(cfq1, cfq7, cfq8)
dim(cfq_all) # 5306   26

# create overall DSST score variable from CFDRIGHT (1999-2000) and CFDDS (2011-2014)
cfq_all <- cfq_all %>%
  mutate(DSST = coalesce(CFDRIGHT, CFDDS))

summary(cfq_all$DSST) 
# range 0 - 117, 875 NAs

# pull just DSST variable to merge
dsst <- cfq_all %>%
  select(SEQN, DSST)

nhanes_data <- left_join(nhanes_data,
                         dsst,
                         by = 'SEQN')
dim(nhanes_data) # 29896    28
  
```

# Prep covariates

In cognition ExWAS analysis:
- Age: variable RIDAGEYR
- Sex: variable RIAGENDR
- Race/ethnicity: variable RIDRETH1
- Education: variable DMDEDUC2 (age 20+), need to recategorize from 5 to 2 categories
- Smoking status: variables SMQ020 and SMQ040, need to import and recategorize
- Blood cotinine: variable LBXCOT
- Urinary creatinine: variable URXUCR
- Fish consumption: variables total_seafood_score and total_seafood_grp, need to import (created in Mortality EWAS project)

For sensitivity analyses:
- Estimated glomerular filtration rate: variable VNEGFR, need to import
- Body composition: variables BMXBMI and BMXWAIST, undecided which to use at the moment
- Alcohol consumption: variables ALQ101, ALQ120Q, ALQ120U, need to recategorize

## Age
```{r}
# overall
summary(nhanes_data$RIDAGEYR)

# included only
nhanes_data %>%
  filter(include == 1) %>%
  summarize(mean_age = mean(RIDAGEYR),
            median_age = median(RIDAGEYR))
```

## Sex
```{r}
# overall
nhanes_data %>%
  count(RIAGENDR)
# 1 = male, 2 = female

# included only
nhanes_data %>%
  filter(include == 1) %>%
  count(RIAGENDR)
```

## Race/ethnicity
```{r}
# overall
nhanes_data %>%
  count(RIDRETH1)
# 1 = mexican american, 2 = other hispanic, 3 = non hispanic white, 4 = non hispanic black, 5 = other race

# included only
nhanes_data %>%
  filter(include == 1) %>%
  count(RIDRETH1)
```

## Education
```{r}
# DMDEDUC2: age 20+, DMDEDUC3: age <20
# will harmonize both variables in order to cover all ages in the included vs excluded table

# overall
table(nhanes_data$DMDEDUC2, useNA = 'always')
# 1 = less than 9th, 2 = some high school with no diploma, 3 = high school grad/GED, 4 = some college or AA, 5 = college grad or above, 7 = refused, 9 = don't know

table(nhanes_data$DMDEDUC3, useNA = 'always')
# 0 - 12 = grade numbers, 13 = HS graduate, 14 = GED, 15 = more than HS, 55 = less than 5th, 66 = less than 9th, 77 = refused, 99 = don't know

# convert 7 and 9 to NA and create binary variable: HS or above vs. below HS
nhanes_data <- nhanes_data %>%
  mutate(DMDEDUC2 = case_when(DMDEDUC2 %in% c(7, 9) ~ NA,
                              .default = DMDEDUC2),
         DMDEDUC3 = case_when(DMDEDUC3 %in% c(77, 99) ~ NA,
                              .default = DMDEDUC3),
         hs_educ = case_when(DMDEDUC2 %in% c(3:5) | DMDEDUC3 %in% c(13:15) ~ 1,
                             DMDEDUC2 %in% c(1, 2) | DMDEDUC3 %in% c(0:12, 55, 66) ~ 0,
                             .default = NA))

table(nhanes_data$hs_educ, useNA = 'always')
table(nhanes_data$hs_educ, nhanes_data$DMDEDUC2, useNA = 'always')
table(nhanes_data$hs_educ, nhanes_data$DMDEDUC3, useNA = 'always')

```

## Smoking status
```{r}
# Load files from NHANES site - these vars were not included in curated datasets for continuous cycles
# To note, these files only cover adults age 20+

smk1 <- read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/1999-2000/SMQ.XPT') %>%
  select(SEQN, SMQ020, SMQ040)
smk7 <- read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/2011-2012/SMQ_G.XPT') %>%
  select(SEQN, SMQ020, SMQ040)
smk8 <- read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/2013-2014/SMQ_H.XPT') %>%
  select(SEQN, SMQ020, SMQ040)

smk_all <- bind_rows(smk1, smk7, smk8)
dim(smk_all) # 18838     3

nhanes_data <- left_join(nhanes_data, 
                         smk_all,
                         by = 'SEQN')
dim(nhanes_data) # 29896    31

table(nhanes_data$SMQ020, useNA = 'ifany')
# smoked at least 100 cigarettes in life: 1 = yes, 2 = no, 7 = refused, 9 = don't know

table(nhanes_data$SMQ040, useNA = 'ifany')
# current smoking: 1 = every day, 2 = some days, 3 = no, 7 = refused, 9 = don't know

table(nhanes_data$SMQ020, nhanes_data$SMQ040, useNA = 'ifany')

nhanes_data <- nhanes_data %>%
  mutate(smk_status = case_when(SMQ020 == 2 ~ 0, # never smokers
                                SMQ020 == 1 & SMQ040 == 3 ~ 1, # former smokers
                                SMQ020 == 1 & SMQ040 %in% c(1, 2) ~ 2, # current smokers
                                .default = NA))

table(nhanes_data$smk_status, useNA = 'always')

```

## Blood cotinine
```{r}
# overall
summary(nhanes_data$LBXCOT)

# included only
nhanes_data %>%
  filter(include == 1) %>%
  summarize(mean_cotinine = mean(LBXCOT, na.rm = T),
            median_cotinine = median(LBXCOT, na.rm = T))

```

## Urinary creatinine
```{r}
# overall
summary(nhanes_data$URXUCR)

# included only
nhanes_data %>%
  filter(include == 1) %>%
  summarize(mean_creatinine = mean(URXUCR, na.rm = T),
            median_creatinine = median(URXUCR, na.rm = T))
```

## Fish consumption
```{r}
# import
fish_consumption <- readRDS('Data/fish_consumption.RDS')

# pull variables needed
fish_merge <- fish_consumption %>%
  select(SEQN, SDDSRVYR, total_seafood_score, total_seafood_grp)

# merge
nhanes_data <- left_join(nhanes_data,
                         fish_merge,
                         by = c('SEQN', 'SDDSRVYR'))
dim(nhanes_data) # 29896    34

# overall
summary(nhanes_data$total_seafood_score)
table(nhanes_data$total_seafood_grp, useNA = 'always')

# included only
nhanes_data %>%
  filter(include == 1) %>%
  summarize(mean_seafood = mean(total_seafood_score, na.rm = T),
            median_seafood = median(total_seafood_score, na.rm = T))

nhanes_data %>%
  filter(include == 1) %>%
  count(total_seafood_grp)
```

## eGFR
```{r}

# overall
summary(nhanes_data$VNEGFR)

# included only
nhanes_data %>%
  filter(include == 1) %>%
  summarize(mean_egfr = mean(VNEGFR, na.rm = T),
            median_egfr = median(VNEGFR, na.rm = T),
            nmiss = sum(is.na(VNEGFR)))

```

## Waist circumference
```{r}
# overall
summary(nhanes_data$BMXWAIST)

# included only
nhanes_data %>%
  filter(include == 1) %>%
  summarize(mean_wc = mean(BMXWAIST, na.rm = T),
            median_wc = median(BMXWAIST, na.rm = T),
            nmiss = sum(is.na(BMXWAIST))) # 370 missing

```

## Alcohol consumption
```{r}

# variable cleaning/creation adapted from code for Pb/Cd and cognition in NHANES analysis (originally done by Yan/Tara)

# ALQ101: had at least 10 drinks in past year?
table(nhanes_data$ALQ101, useNA = 'always')

nhanes_data %>%
  filter(include == 1) %>%
  count(ALQ101)
# 1 yes, 2 no, 9 don't know; 270 NAs in included grp

# convert 9 to NA
nhanes_data <- nhanes_data %>%
  mutate(ALQ101 = case_when(ALQ101 == 9 ~ NA,
                            .default = ALQ101))

table(nhanes_data$ALQ101, useNA = 'always')

# ALQ120Q: alcohol consumption frequency in past 12 months; ALQ120U: units to match ALQ120Q 
table(nhanes_data$ALQ120Q, useNA = 'always')
# convert ALQ120Q = 777 and 999 to NA
# create new number of drinks/12 months var: if ALQ101 = 2 (no), then set to 0 drinks - otherwise ALQ120Q
nhanes_data <- nhanes_data %>%
  mutate(ALQ120Q = case_when(ALQ120Q %in% c(777, 999) ~ NA,
                             .default = ALQ120Q),
         drinks_past_yr = case_when(ALQ101 == 2 ~ 0,
                                    .default = ALQ120Q))

# based on units from ALQ120U, create multiplier to convert to drinks per month
table(nhanes_data$ALQ120U, useNA = 'always')
# 1 = week, 2 = month, 3 = year
# if week: multiply by 52weeks/12months = 4.33
# if month: multiply by 1
# if year: multiply by 1year/12months = 0.083

nhanes_data <- nhanes_data %>%
  mutate(alc_unit_mult = case_when(ALQ120U == 1 ~ 4.33,
                                   ALQ120U == 2 ~ 1,
                                   ALQ120U == 3 ~ 0.083,
                                   drinks_past_yr == 0 ~ 1)) # fill in a 1 if no alc consumed, otherwise it would try to mult by NA

table(nhanes_data$alc_unit_mult, useNA = 'always')

# calculate number of drinks per month
nhanes_data <- nhanes_data %>%
  mutate(alc_month_freq = drinks_past_yr * alc_unit_mult)

summary(nhanes_data$alc_month_freq, useNA = 'always')

# confirm that no NAs in participants with 0 drinks
alq_0_drinks <- nhanes_data %>%
  filter(drinks_past_yr == 0)

summary(alq_0_drinks$alc_month_freq) # all 0, good

# create binary alcohol intake (0 if <=4/month, 1 if > 4/month)
nhanes_data <- nhanes_data %>%
  mutate(alc4 = case_when(alc_month_freq <= 4 ~ 0,
                          alc_month_freq > 4 ~ 1))

table(nhanes_data$alc4, useNA = 'always')

# check availability in included group
nhanes_data %>%
  filter(include == 1) %>%
  count(alc4) # 280 NAs, not too excessive

```

# Social Vulnerability Index - will only be used in the interaction analysis, not main effect EWAS

Components:
1. Income - variable included (INDFMPIR)
2. Education - variable included (DMDEDUC2)
3. Race/ethnicity - variable included (RIDRETH1)
4. Health insurance - need to import
5. Housing tenure - need to import
6. Food security - need to import
7. Employment - need to import

## Income
```{r}
summary(nhanes_data$INDFMPIR)
```

## Education
```{r}
table(nhanes_data$hs_educ, useNA = 'always')
```

## Race/ethnicity
```{r}
table(nhanes_data$RIDRETH1, useNA = 'always')
```

## Health insurance
```{r}
# Note that questions/variables are different between 1999-2000 and 2011-2014
hiq1 = read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/1999-2000/HIQ.XPT')
hiq7 = read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/2011-2012/HIQ_G.XPT')
hiq8 = read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/2013-2014/HIQ_H.XPT')

# clean 1999-2000 first
hiq1 <- hiq1 %>%
  select(SEQN, HID010, HID030A, HID030B, HID030C, HID030D)

table(hiq1$HID010, useNA = 'always')
# 1 = yes ins, 2 = no ins, 7/9/NA
# 2038 uninsured

table(hiq1$HID030A, useNA = 'always') # 5215 yes private ins
table(hiq1$HID030B, useNA = 'always') # 1232 yes medicare
table(hiq1$HID030C, useNA = 'always') # 1492 yes medicaid or chip
table(hiq1$HID030D, useNA = 'always') # 534 yes other govt ins

# yes (1) to HID030A = private
# yes (1) to B, C, or D = govt
# participants can answer Y to more than 1 type - deciding to prioritize Y to private over govt types for categorization

hiq1 <- hiq1 %>%
  mutate(ins_status = case_when(HID010 == 2 ~ 1, # no ins
                                HID030A == 1 ~ 0, # private ins
                                (HID030B == 1 | HID030C == 1 | HID030D == 1) ~ 0.5)) # govt ins

table(hiq1$ins_status, useNA = 'always')
# 266 NAs

# 2011-2014 next
hiq_7_8 = rbind(hiq7, hiq8) %>%
    select(SEQN, HIQ011, HIQ031A, HIQ031B, HIQ031C, HIQ031D, HIQ031E, HIQ031F, HIQ031G, HIQ031H, HIQ031I) 

table(hiq_7_8$HIQ011, useNA = 'always') # no insurance (2) = 3273
table(hiq_7_8$HIQ031A, useNA = 'always') # private ins (14) = 9154
table(hiq_7_8$HIQ031B, useNA = 'always') # medicare (15) = 2432
table(hiq_7_8$HIQ031C, useNA = 'always') # medi-gap (16) = 25
table(hiq_7_8$HIQ031D, useNA = 'always') # medicaid (17) = 4577
table(hiq_7_8$HIQ031E, useNA = 'always') # chip (18) = 151
table(hiq_7_8$HIQ031F, useNA = 'always') # military (19) = 402
table(hiq_7_8$HIQ031G, useNA = 'always') # indian health service (20) = 21
table(hiq_7_8$HIQ031H, useNA = 'always') # state sponsored (21) = 1155
table(hiq_7_8$HIQ031I, useNA = 'always') # other govt (22) = 505

# from documentation of 05-06 questionnaire https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/HIQ_D.htm (old = new):
# HID010 = HIQ011
# private ins HID030A = HIQ031A (private) and HIQ031C (medigap)
# govt ins HID030B, HID030C, HID030D = HIQ031B, HIQ031D thru I (covers medicare, medicaid, chip, military, IHS, state sponsored, and other govt ins)

hiq_7_8 <- hiq_7_8 %>%
  mutate(ins_status = case_when(HIQ011 == 2 ~ 1, # no ins
                                (HIQ031A == 14 | HIQ031C == 16) ~ 0, # private ins
                                (HIQ031B == 15 | HIQ031D == 17 | HIQ031E == 18 | HIQ031F == 19 | 
                                   HIQ031G == 20 | HIQ031H == 21 | HIQ031I == 22) ~ 0.5)) # govt ins

table(hiq_7_8$ins_status, useNA = 'always') # 116 NAs

# combine the two sets
hiq1_keep <- hiq1 %>%
  select(SEQN, ins_status)
hiq_7_8_keep <- hiq_7_8 %>%
  select(SEQN, ins_status)

hiq_all <- bind_rows(hiq1_keep, hiq_7_8_keep)

table(hiq_all$ins_status, useNA = 'always')

nhanes_data <- left_join(nhanes_data,
                         hiq_all,
                         by = 'SEQN')

table(nhanes_data$ins_status, useNA = 'always')

```

## Housing tenure
```{r}
# Var needed: HOQ065
hoq1 = read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/1999-2000/HOQ.XPT')
hoq7 = read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/2011-2012/HOQ_G.XPT')
hoq8 = read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/2013-2014/HOQ_H.XPT')

# HOQ065 = housing tenure, own/rent/other
hoq_all = bind_rows(hoq1, hoq7, hoq8) %>%
  select(SEQN, HOQ065)

nhanes_data = left_join(nhanes_data,
                        hoq_all,
                        by = 'SEQN')

table(nhanes_data$HOQ065, useNA = 'always')

# convert 7 and 9 to NA
nhanes_data <- nhanes_data %>%
  mutate(HOQ065 = case_when(HOQ065 == 7 ~ NA,
                            HOQ065 == 9 ~ NA,
                            .default = HOQ065))

table(nhanes_data$HOQ065, useNA = 'always')
```

## Food security
```{r}
# Vars needed: ADFDSEC and FSDAD
fsq1 <- read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/1999-2000/FSQ.XPT')
fsq7 <- read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/2011-2012/FSQ_G.XPT')
fsq8 <- read_xpt('https://wwwn.cdc.gov/Nchs/Nhanes/2013-2014/FSQ_H.XPT')

fsq_all <- bind_rows(fsq1, fsq7, fsq8) %>%
  select(SEQN, ADFDSEC, FSDAD)
# 1999-2000 uses ADFDSEC, 2011-2014 uses FSDAD

# merge into sample dataset
nhanes_data <- left_join(nhanes_data,
                         fsq_all,
                         by = 'SEQN')

# harmonize variables across cycles
nhanes_data <- nhanes_data %>%
  mutate(ad_foodsec = coalesce(ADFDSEC, FSDAD))
# 1 = full security, 2 = marginal, 3 = low, 4 = very low

table(nhanes_data$ad_foodsec, useNA = 'always')

```

## Employment
```{r}
# Vars needed: OCD150 and OCQ380

table(nhanes_data$OCD150, useNA = 'always')
table(nhanes_data$OCQ380, useNA = 'always')

# create binary variable for employment status
nhanes_data <- nhanes_data %>%
  mutate(employed = case_when(OCD150 %in% c(1, 2) ~ 1, # employed
                              OCD150 == 3 | OCQ380 %in% c(1:7) ~ 0, # not employed (incl. retired)
                              .default = NA))

table(nhanes_data$employed, useNA = 'always')
table((nhanes_data %>% filter(include == 1))$employed, useNA = 'always')

```

## Create SVI score
```{r}
# check variables that are used to calculate components
summary(nhanes_data$INDFMPIR) # 1.5 or below = 1 point
table(nhanes_data$DMDEDUC2, useNA = 'always') # less than HS (1 or 2) = 1 point
table(nhanes_data$HOQ065, useNA = 'always') # rent or other (2, 3) = 1 point
table(nhanes_data$RIDRETH1, useNA = 'always') # not non hispanic white (3) = 1 point
table(nhanes_data$ins_status, useNA = 'always') # govt ins = 0.5 pt, no ins = 1 point
table(nhanes_data$ad_foodsec, useNA = 'always') # less than full security (2, 3, or 4) = 1 point
table(nhanes_data$employed, useNA = 'always') # unemployed (0) = 1 point

nhanes_data = nhanes_data %>%
# Income
# 1 point on index if 1.5 or below, 0 if above
    mutate(svi_income = ifelse(INDFMPIR <= 1.5, 1, 0)) %>%

# Education
# 1 point on index if no hs ed, 0 if yes
    mutate(svi_educ = ifelse(DMDEDUC2 < 3, 1, 0)) %>%

# Health insurance
# Coding is same as ins_status: 1 point on index if no ins, 0.5 if govt ins, 0 if private ins
    mutate(svi_ins = ins_status) %>%

# Race/ethnicity
# Coding: 0 points on index if non hispanic white, 1 point for all others
    mutate(svi_race = ifelse(RIDRETH1 == 3, 0, 1)) %>%

# Housing tenure
# 1 point for rent or other, 0 for own home
    mutate(svi_tenure = case_when(HOQ065 == 1 ~ 0, # homeowner
                                  HOQ065 %in% c(2, 3) ~ 1, # rent or other
                                  .default = NA)) %>%
  
# Food security
# 1 point on index if less than full security, 0 if fully secure
    mutate(svi_foodsec = ifelse(ad_foodsec > 1, 1, 0)) %>%
  
# Employment
# 1 point if unemployed or retired, 0 if employed
    mutate(svi_employment = ifelse(employed == 0, 1, 0)) %>%

# sum up for SVI overall
    mutate(svi = svi_income + svi_educ + svi_ins + svi_race + 
             svi_tenure + svi_foodsec + svi_employment)

table(nhanes_data$svi_income, useNA = 'always')
table(nhanes_data$svi_educ, nhanes_data$DMDEDUC2, useNA = 'always')
table(nhanes_data$svi_ins, nhanes_data$ins_status, useNA = 'always')
table(nhanes_data$svi_race, nhanes_data$RIDRETH1, useNA = 'always')
table(nhanes_data$svi_tenure, nhanes_data$HOQ065, useNA = 'always')
table(nhanes_data$svi_foodsec, nhanes_data$ad_foodsec, useNA = 'always')
table(nhanes_data$svi_employment, nhanes_data$employed, useNA = 'always')
table(nhanes_data$svi, useNA = 'always')

summary(nhanes_data$svi) 

nhanes_data %>%
  filter(include == 1) %>%
  summarize(mean = mean(svi, na.rm = T),
            nmiss = sum(is.na(svi)),
            n = n() - nmiss)
# in analysis group, 621 missing SVI overall (at least 1 component missing -> SVI missing)

```

# Remove unnecessary variables from nhanes_data
```{r}

dim(nhanes_data) # 29896    52
colnames(nhanes_data)

cognition_ewas_svi_to_impute <- nhanes_data %>%
  select(-c('SEQN_new', 'OCD150', 'OCQ380', 'ALQ101', 'ALQ120Q', 'ALQ120U', 'aged_60plus',
            'SMQ020', 'SMQ040', 'drinks_past_yr', 'alc_unit_mult', 'ADFDSEC', 'FSDAD'))
dim(cognition_ewas_svi_to_impute) # 29896    39

```

# Add variable labels
```{r}

get_label(cognition_ewas_svi_to_impute)

# fill in blank variable labels
cognition_ewas_svi_to_impute <- cognition_ewas_svi_to_impute %>%
  var_labels(include = 'Included in analysis',
             DSST = 'DSST score (all cycles)',
             hs_educ = 'Education (high school completion)',
             smk_status = 'Smoking status',
             alc_month_freq = 'Alcohol consumption (drinks per month)',
             alc4 = 'Alcohol consumption (Over 4 drinks per month)',
             ins_status = 'Health insurance coverage type',
             HOQ065 = 'Home owned, bought, rented, or other',
             ad_foodsec = 'Food security level (all cycles)',
             employed = 'Employment status',
             svi_income = 'SVI income component',
             svi_educ = 'SVI education component',
             svi_ins = 'SVI health insurance component',
             svi_race = 'SVI race/ethnicity component',
             svi_tenure = 'SVI housing tenure component',
             svi_foodsec = 'SVI food security component',
             svi_employment = 'SVI employment component',
             svi = 'Social Vulnerability Index (SVI) score')

# add value labels for categorical variables
cognition_ewas_svi_to_impute <- cognition_ewas_svi_to_impute %>%
  val_labels(SDDSRVYR = c('1999-2000', 
                          '2011-2012', 
                          '2013-2014'),
             RIAGENDR = c('Male',
                          'Female'),
             RIDRETH1 = c('Mexican American',
                          'Other Hispanic',
                          'Non-Hispanic White',
                          'Non-Hispanic Black',
                          'Other Race'),
             include = c('Excluded',
                         'Included'),
             hs_educ = c('Did not complete high school',
                         'Completed high school or above'),
             smk_status = c('Never smoker',
                            'Former smoker',
                            'Current smoker'),
             alc4 = c('0-4 drinks per month',
                      '>4 drinks per month'),
             ins_status = c('Private insurance',
                            'Government insurance',
                            'No insurance'),
             ad_foodsec = c('Full security',
                            'Marginal security',
                            'Low security',
                            'Very low security'),
             employed = c('Unemployed or retired',
                          'Employed'),
             svi_income = c('PIR > 1.5', 
                            'PIR <= 1.5'), 
             svi_educ = c('Completed high school or above',
                          'Did not complete high school'), 
             svi_ins = c('Private insurance', 
                         'Government insurance', 
                         'No insurance'), 
             svi_race = c('Non-Hispanic White', 
                          'Not Non-Hispanic White'), 
             svi_tenure = c('Homeowner', 
                            'Renter or other'),
             svi_foodsec = c('Full security',
                             'Less than full security'),
             svi_employment = c('Employed',
                                'Unemployed or retired'))

str(cognition_ewas_svi_to_impute)

```

# Save dataset to be imputed
```{r}

saveRDS(cognition_ewas_svi_to_impute, file = 'Data/cognition_ewas_svi_to_impute.RDS')

```


