---
title: "Descriptive statistics tables: cognition ExWAS"
author: "Erika Walker"
date: "`r Sys.Date()`"
output: pdf_document
---

Purpose: Create descriptive statistics tables (both survey weighted and unweighted) for cognition ExWAS analysis

Presented demographic tables will be weighted, except for unweighted frequencies (ex: N missing; subgroup sample sizes)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gtsummary)
library(flextable)
library(tidyverse)
library(sjlabelled)
library(survey)
library(gt)
```

# Load data 
```{r}
ewas_data <- readRDS('Data/cognition_ewas_svi_to_impute.RDS')
# N = 29896 (3 full cycles)

#summary(ewas_data)

```

# Descriptives by cycle
Comparing cycles 1999-2000 vs 2011-2014
Unweighted
```{r}
# create categories for the cycle periods for comparison and use the categorical var labels
tbl1_data <- ewas_data %>%
  mutate(cycles = ifelse(SDDSRVYR == 1, "1999-2000", "2011-2014")) %>%
  as_label(RIAGENDR, RIDRETH1, hs_educ, smk_status, alc4) %>%
  mutate(normal_kidney = case_when(VNEGFR >= 60 ~ 'eGFR >= 60',
                                   VNEGFR < 60 ~ 'eGFR < 60')) %>%
  mutate(total_seafood_grp = factor(total_seafood_grp,
                                    labels = c('0 times',
                                               '1-3 times',
                                               '4+ times')))

table(tbl1_data$total_seafood_grp)

# generate table 1 unweighted descriptive statistics by cycle
tbl_1 <- tbl1_data %>%
  filter(include == 1) %>%
  tbl_summary(by = 'cycles',
              include = c(DSST, RIDAGEYR, RIAGENDR, RIDRETH1, hs_educ, smk_status, LBXCOT, alc4, 
                          BMXWAIST, URXUCR, total_seafood_grp, normal_kidney),
              label = c(DSST ~ 'Digit Symbol Substitution Test (DSST)',
                        RIDAGEYR ~ 'Age (years)',
                        RIAGENDR ~ 'Sex',
                        RIDRETH1 ~ 'Race/ethnicity',
                        hs_educ ~ 'Education',
                        smk_status ~ 'Smoking status',
                        LBXCOT ~ 'Serum cotinine (ng/mL)',
                        alc4 ~ 'Alcohol consumption',
                        BMXWAIST ~ 'Waist circumference (cm)',
                        URXUCR ~ 'Urinary creatinine (mg/dL)',
                        total_seafood_grp ~ 'Fish and seafood consumption in past 30 days',
                        normal_kidney ~ 'Estimated glomerular filtration rate (eGFR; ml/min/1.73 m2)'),
              missing_text = 'Missing',
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(all_continuous() ~ c(2, 2),
                            all_categorical() ~ c(0, 2))) %>%
  add_overall() %>%
  add_p(list(all_continuous() ~ 't.test',
             all_categorical() ~ 'chisq.test')) %>%
  bold_labels() %>%
  modify_header(label = '**Variable**',
                stat_0 = '**Overall (N = 4,982)**',
                stat_1 = '**1999-2000 (N = 1,562)**',
                stat_2 = '**2011-2014 (N = 3,420)**') %>%
  modify_caption('Table 1. Unweighted descriptive statistics of participants aged 60+ in the National Health and Nutrition Examination Survey (NHANES), stratified by cycle of participation (N = 4,982).')

tbl_1

tbl_1 %>%
  as_flex_table() %>%
  save_as_docx(path = 'Output Cog ExWAS/Cog ExWAS unweighted descriptive stats by cycle May 31 2024.docx')

```

# Weighted descriptives by cycle
```{r}
# adjust weights
tbl1_data <- tbl1_data %>%
    mutate(adjusted_weights = (1/3)*WTMEC2YR)

# create survey object
tbl1_wt_data <- svydesign(ids = ~SDMVPSU, 
                          strata = ~SDMVSTRA,
                          weights = ~adjusted_weights,
                          nest = TRUE,
                          data = tbl1_data)

# subset to included participants
tbl1_wt_subset <- subset(tbl1_wt_data,
                         include == 1)

tbl_1_wt <- tbl_svysummary(data = tbl1_wt_subset, 
            by = 'cycles', 
            include = c(DSST, RIDAGEYR, RIAGENDR, RIDRETH1, hs_educ, smk_status, LBXCOT, alc4, 
                          BMXWAIST, URXUCR, total_seafood_grp, normal_kidney),
              label = c(DSST ~ 'Digit Symbol Substitution Test (DSST)',
                        RIDAGEYR ~ 'Age (years)',
                        RIAGENDR ~ 'Sex',
                        RIDRETH1 ~ 'Race/ethnicity',
                        hs_educ ~ 'Education',
                        smk_status ~ 'Smoking status',
                        LBXCOT ~ 'Serum cotinine (ng/mL)',
                        alc4 ~ 'Alcohol consumption',
                        BMXWAIST ~ 'Waist circumference (cm)',
                        URXUCR ~ 'Urinary creatinine (mg/dL)',
                        total_seafood_grp ~ 'Fish and seafood consumption in past 30 days',
                        normal_kidney ~ 'Estimated glomerular filtration rate (eGFR; ml/min/1.73 m2)'),
            missing_text = 'Missing',
            statistic = list(all_continuous() ~ "{mean} ({sd})",
                             all_categorical() ~ "{p}%"),
            digits = list(all_continuous() ~ c(2, 2),
                          all_categorical() ~ 2)) %>%
  add_overall() %>%
  add_p(list(all_continuous() ~ 'svy.t.test',
             all_categorical() ~ 'svy.chisq.test')) %>%
  bold_labels() %>%
  modify_header(label = '**Variable**',
                stat_0 = '**Overall (N = 4,982)**',
                stat_1 = '**1999-2000 (N = 1,562)**',
                stat_2 = '**2011-2014 (N = 3,420)**') %>%
  modify_caption('Table 1. Weighted descriptive statistics of participants aged 60+ in the National Health and Nutrition Examination Survey (NHANES), stratified by cycle of participation (N = 4,982).')

tbl_1_wt

tbl_1_wt %>%
  as_flex_table() %>%
  save_as_docx(path = 'Output Cog ExWAS/Cog ExWAS weighted descriptive stats by cycle May 31 2024.docx')

```


# Descriptives by cognitive status
Mild cognitive impairment dichotomized (cutoff = 25th pctl)
25th pctl in imputed set: 29
25th pctl in unimputed set: 31
Going with 29 as cutoff so it matches the models using the imp dataset
```{r}
summary((ewas_data %>% filter(include == 1))$DSST) 

#mci_cutoff <- as.numeric(summary((ewas_data %>% filter(include == 1))$DSST))[2]

mci_cutoff <- 29

tbl2_data <- ewas_data %>%
  mutate(mci = ifelse(DSST < mci_cutoff, 'Mild cognitive impairment', 'No cognitive impairment')) %>%
  as_label(RIAGENDR, RIDRETH1, hs_educ, smk_status, alc4, SDDSRVYR) %>%
  mutate(normal_kidney = case_when(VNEGFR >= 60 ~ 'eGFR >= 60',
                                   VNEGFR < 60 ~ 'eGFR < 60')) %>%
  mutate(total_seafood_grp = factor(total_seafood_grp,
                                    labels = c('0 times',
                                               '1-3 times',
                                               '4+ times')))

table(tbl2_data$mci)

# force NA to category for mci so it can show as a column
tbl2_data <- tbl2_data %>%
  mutate(mci = fct_na_value_to_level(mci,
                                     level = 'Missing DSST score'))

table(tbl2_data$mci)
table((tbl2_data %>% filter(include == 1))$mci)

tbl_2 <- tbl2_data %>%
  filter(include == 1) %>%
  tbl_summary(by = 'mci', 
              include = c(DSST, RIDAGEYR, RIAGENDR, RIDRETH1, hs_educ, smk_status, LBXCOT, alc4, 
                          BMXWAIST, URXUCR, total_seafood_grp, normal_kidney, SDDSRVYR),
              label = c(DSST ~ 'Digit Symbol Substitution Test (DSST)',
                        RIDAGEYR ~ 'Age (years)',
                        RIAGENDR ~ 'Sex',
                        RIDRETH1 ~ 'Race/ethnicity',
                        hs_educ ~ 'Education',
                        smk_status ~ 'Smoking status',
                        LBXCOT ~ 'Serum cotinine (ng/mL)',
                        alc4 ~ 'Alcohol consumption',
                        BMXWAIST ~ 'Waist circumference (cm)',
                        URXUCR ~ 'Urinary creatinine (mg/dL)',
                        total_seafood_grp ~ 'Fish and seafood consumption in past 30 days',
                        normal_kidney ~ 'Estimated glomerular filtration rate (eGFR; ml/min/1.73 m2)',
                        SDDSRVYR ~ 'NHANES cycle'),
            missing_text = 'Missing',
            statistic = list(all_continuous() ~ "{mean} ({sd})",
                             all_categorical() ~ "{n} ({p}%)"),
            digits = list(all_continuous() ~ c(2, 2),
                          all_categorical() ~ c(0, 2))) %>%
  add_p(list(all_continuous() ~ 'aov',
             all_categorical() ~ 'chisq.test')) %>%
  bold_labels() %>%
  modify_header(label = '**Variable**',
                stat_1 = '**Mild cognitive impairment (N = 889)**',
                stat_2 = '**No cognitive impairment (N = 3,372)**',
                stat_3 = '**Missing DSST score (N = 721)**') %>%
  modify_caption('Table 2. Unweighted descriptive statistics of participants aged 60+ in the National Health and Nutrition Examination Survey (NHANES) years 1999-2000 and 2011-2014, stratified by cognitive status determined via Digit Symbol Substitution Test (DSST) score (N = 4,982).')
  
tbl_2

tbl_2 %>%
  as_flex_table() %>%
  save_as_docx(path = 'Output Cog ExWAS/Cog ExWAS unweighted descriptive stats by cognitive status May 31 2024.docx')

```

# Weighted descriptives by cognitive status
```{r}
# adjust weights
tbl2_data <- tbl2_data %>%
    mutate(adjusted_weights = (1/3)*WTMEC2YR)

tbl2_wt_data <- svydesign(ids = ~SDMVPSU, 
                          strata = ~SDMVSTRA,
                          weights = ~adjusted_weights,
                          nest = TRUE,
                          data = tbl2_data)

tbl2_wt_subset <- subset(tbl2_wt_data,
                         include == 1)

tbl_2_wt <- tbl_svysummary(data = tbl2_wt_subset, 
            by = "mci", 
            include = c(DSST, RIDAGEYR, RIAGENDR, RIDRETH1, hs_educ, smk_status, LBXCOT, alc4, 
                          BMXWAIST, URXUCR, total_seafood_grp, normal_kidney, SDDSRVYR),
              label = c(DSST ~ 'Digit Symbol Substitution Test (DSST)',
                        RIDAGEYR ~ 'Age (years)',
                        RIAGENDR ~ 'Sex',
                        RIDRETH1 ~ 'Race/ethnicity',
                        hs_educ ~ 'Education',
                        smk_status ~ 'Smoking status',
                        LBXCOT ~ 'Serum cotinine (ng/mL)',
                        alc4 ~ 'Alcohol consumption',
                        BMXWAIST ~ 'Waist circumference (cm)',
                        URXUCR ~ 'Urinary creatinine (mg/dL)',
                        total_seafood_grp ~ 'Fish and seafood consumption in past 30 days',
                        normal_kidney ~ 'Estimated glomerular filtration rate (eGFR; ml/min/1.73 m2)',
                        SDDSRVYR ~ 'NHANES cycle'),
            missing_text = 'Missing',
            statistic = list(all_continuous() ~ "{mean} ({sd})",
                             all_categorical() ~ "{p}%"),
            digits = list(all_continuous() ~ c(2, 2),
                          all_categorical() ~ 2)) %>%
  add_p(list(all_continuous() ~ 'svy.kruskal.test',
             all_categorical() ~ 'svy.chisq.test')) %>%
  bold_labels() %>%
  modify_header(label = '**Variable**',
                stat_1 = '**Mild cognitive impairment (N = 889)**',
                stat_2 = '**No cognitive impairment (N = 3,372)**',
                stat_3 = '**Missing DSST score (N = 721)**') %>%
  modify_caption('Table 2. Weighted descriptive statistics of participants aged 60+ in the National Health and Nutrition Examination Survey (NHANES) years 1999-2000 and 2011-2014, stratified by cognitive status determined via Digit Symbol Substitution Test (DSST) score (N = 4,982).')
  
tbl_2_wt

tbl_2_wt %>%
  as_flex_table() %>%
  save_as_docx(path = 'Output Cog ExWAS/Cog ExWAS weighted descriptive stats by cognitive status May 31 2024.docx')
```

# Chemical descriptives - weighted
Now weighted, Jul 9
```{r}
chems <- read_csv('Data/chemicals_clean.csv', na = '.') %>%
  filter(SDDSRVYR %in% c(1, 7, 8))

weights <- read_csv('Data/weights_clean.csv', na = '.') %>%
  filter(SDDSRVYR %in% c(1, 7, 8))

dictionary <- read_csv('Data/use_these_chems.csv')

chem_list <- dictionary$variable_codename_use

wt_list <- paste0('WT_', chem_list)

# also add weight list to dictionary
dictionary$weight_variable <- paste0('WT_', dictionary$variable_codename_use)

# reduce chems and weights files to included vars only and pivot
chems_long <- chems %>%
  select(SEQN, SDDSRVYR, all_of(chem_list)) %>%
  pivot_longer(cols = all_of(chem_list),
               names_to = 'chem_name',
               values_to = 'chem_value')

weights_long <- weights %>%
  select(SEQN, SDDSRVYR, all_of(wt_list)) %>%
  pivot_longer(cols = all_of(wt_list),
               names_to = 'weight_name',
               values_to = 'weight_value')

# merge the long datasets
chems_weights_long <- left_join(weights_long,
                                dictionary,
                                join_by(weight_name == weight_variable)) %>%
  select(SEQN, SDDSRVYR, weight_name, weight_value, variable_codename_use) %>%
  left_join(.,
            chems_long,
            join_by(SEQN == SEQN,
                    SDDSRVYR == SDDSRVYR,
                    variable_codename_use == chem_name)) %>%
  rename(chem_name = variable_codename_use)

# pull number of cycles per chem
cycles <- chems_long %>%
  drop_na(chem_value) %>%
  group_by(chem_name) %>%
  count(SDDSRVYR) %>%
  group_by(chem_name) %>%
  mutate(num_cycles = length(SDDSRVYR)) %>%
  select(-n, -SDDSRVYR) %>%
  unique()

chems_weights_long <- left_join(chems_weights_long,
                                cycles)

# adjust weight
chems_weights_long <- chems_weights_long %>%
  mutate(weight_adj = (1/num_cycles)*weight_value)

# add survey design vars
svy_vars <- ewas_data %>%
  select(SEQN, SDDSRVYR, SDMVPSU, SDMVSTRA, include)

chems_weights_long <- left_join(chems_weights_long,
                                svy_vars)

# function for weighted stats
desc_stats_wt <- function(chem_subset) {
  
  # remove missing weights
  chem_subset <- chem_subset %>%
    drop_na(weight_adj)
  
  # set up survey weighting
  chem_subset_svy <- svydesign(ids = ~SDMVPSU, 
                               strata = ~SDMVSTRA,
                               weights = ~weight_adj,
                               nest = TRUE,
                               data = chem_subset)
  
  # subset to included participants
  chem_subset_svy_incl <- subset(chem_subset_svy,
                                 include == 1)
  
  # summary stats
  unwt_n <- chem_subset %>%
    filter(include == 1) %>%
    drop_na(chem_value) %>%
    nrow()
  wt_mean <- svymean(~chem_value, design = chem_subset_svy_incl, na.rm = T)[1]
  wt_sd <- sqrt(svyvar(~chem_value, design = chem_subset_svy_incl, na.rm = T))[1]
  wt_geom_mean <- exp(svymean(~log(chem_value), design = chem_subset_svy_incl, na.rm = T))[1]
  wt_geom_sd <- exp(sqrt(svyvar(~log(chem_value), design = chem_subset_svy_incl, na.rm = T)))[1]
  unwt_min <- min((chem_subset %>% filter(include == 1))$chem_value, na.rm = T)
  wt_p25 <- svyquantile(~chem_value, design = chem_subset_svy_incl, quantiles = 0.25)$chem_value[1]
  wt_med <- svyquantile(~chem_value, design = chem_subset_svy_incl, quantiles = 0.5)$chem_value[1]
  wt_p75 <- svyquantile(~chem_value, design = chem_subset_svy_incl, quantiles = 0.75)$chem_value[1]
  unwt_max <- max((chem_subset %>% filter(include == 1))$chem_value, na.rm = T)
  
  tbl_stats <- bind_cols(chem_name = unique(chem_subset$chem_name),
                         unwt_n = unwt_n,
                         wt_mean = wt_mean,
                         wt_sd = wt_sd,
                         wt_geom_mean = wt_geom_mean,
                         wt_geom_sd = wt_geom_sd,
                         unwt_min = unwt_min,
                         wt_p25 = wt_p25,
                         wt_med = wt_med,
                         wt_p75 = wt_p75,
                         unwt_max = unwt_max)
  
  return(tbl_stats)
  
}

# generate desc stats
chem_desc_weighted <- chems_weights_long %>%
  group_by(chem_name) %>%
  do(desc_stats_wt(.))

# merge chem dictionary info for more detail
chem_desc_weighted <- left_join(chem_desc_weighted,
                                dictionary,
                                join_by(chem_name == variable_codename_use,
                                        unwt_n == n_measures)) %>%
  select(-c('unique_cycles', 'comment_codename_use', 'weight_variable')) 

# reorganize columns
chem_desc_weighted <- chem_desc_weighted %>%
  relocate(variable_description_use, .after = chem_name) %>%
  relocate(chemical_family, .after = pct_above_lod) %>%
  relocate(unwt_n, .after = unwt_max)

# save
write_csv(chem_desc_weighted, file = 'Output Cog ExWAS/Chem descriptive stats weighted Jul 9 2024.csv')

```

# Chemical descriptives - unweighted
```{r}

# reduce chems file to included participants and chems
included_seqns <- ewas_data %>%
  filter(include == 1) %>%
  pull(SEQN)

chems_for_unwt <- chems %>%
  filter(SEQN %in% included_seqns) %>%
  select(SEQN, SDDSRVYR, all_of(chem_list))

# summary stats table
chem_desc <- chems_for_unwt %>%
  pivot_longer(cols = all_of(chem_list),
               values_to = 'measure',
               names_to = 'chem_name') %>%
  drop_na(measure) %>%
  group_by(chem_name) %>%
  summarize(n = n(),
            mean = mean(measure), 
            sd = sd(measure), 
            geometric_mean = exp(mean(log(measure))),
            geometric_sd = exp(sd(log(measure))),
            min = min(measure), 
            p25 = quantile(measure, prob = 0.25),
            med = median(measure),
            p75 = quantile(measure, prob = 0.75),
            max = max(measure))

# merge chem dictionary info for more detail
chem_desc <- left_join(chem_desc,
                       dictionary,
                       join_by(chem_name == variable_codename_use,
                               n == n_measures)) %>%
  select(-comment_codename_use, -unique_cycles) 

# reorganize columns
chem_desc <- chem_desc %>%
  relocate(variable_description_use, .after = chem_name) %>%
  relocate(chemical_family, .after = pct_above_lod) %>%
  relocate(n, .after = max)
  
# save
#write_csv(chem_desc, file = 'Output Cog ExWAS/Chem descriptive stats May 15 2024.csv')

```

# Imputation summary
```{r}
# list of variables that I imputed
vars_to_impute <- c('INDFMPIR', 'DMDEDUC2', 'URXUCR', 'BMXBMI', 'BMXWAIST', 'LBXCOT', 'DSST', 
                    'smk_status', 'total_seafood_score', 'alc_month_freq', 'ins_status', 
                    'HOQ065', 'ad_foodsec', 'employed')
                    

# pull included participants only
included <- ewas_data %>%
  filter(include == 1)

# create a table of the missing Ns (for included participants only)
missings <- as.data.frame(sapply(included, function(x) sum(is.na(x)))) %>%
  filter(rownames(.) %in% vars_to_impute) %>%
  rownames_to_column(var = 'variable')

# rename N column
colnames(missings) <- c('variable', 'n_missing')

# add column for imputation method used
# I'm pulling this info manually from the imputation code file

missings$variable

missings$imp_method <- c('Predictive mean modeling', 'Proportional odds modeling', 'Predictive mean modeling', 
                         'Predictive mean modeling', 'Predictive mean modeling', 'Predictive mean modeling', 
                         'Predictive mean modeling', 'Polytomous logistic regression', 'Predictive mean modeling', 
                         'Predictive mean modeling', 'Polytomous logistic regression', 'Polytomous logistic regression', 
                         'Proportional odds model', 'Logistic regression')

# add name of variable
missings$description <- get_label(ewas_data %>% select(all_of(vars_to_impute)))

# clean up some of these
missings$description[1] <- 'Poverty income ratio (PIR)'
missings$description[2] <- 'Education level'
missings$description[7] <- 'DSST score'
missings$description[12] <- 'Housing tenure'
missings$description[13] <- 'Food security level'

# make gt table
missings_gt <- missings %>%
  select(-variable) %>%
  relocate(description, .before = n_missing) %>%
  gt() %>%
  cols_label(description = 'Variable',
             n_missing = 'N missing',
             imp_method = 'Imputation method') %>%
  tab_style(style = cell_text(weight = 'bold'), 
            locations = cells_column_labels()) %>%
  tab_style(style = cell_text(align = 'center'), 
          locations = cells_column_labels()) %>%
  cols_align(align = 'left')

missings_gt

#gtsave(missings_gt, filename = 'Output Cog ExWAS/Imputation info table May 10 2024.docx')

# overall % with at least one missing var
n_miss <- included %>%
  mutate(educ_m = is.na(DMDEDUC2),
         creat_m = is.na(URXUCR),
         waist_m = is.na(BMXWAIST),
         cot_m = is.na(LBXCOT),
         dsst_m = is.na(DSST),
         smk_m = is.na(smk_status),
         seafood_m = is.na(total_seafood_score),
         alc_m = is.na(alc_month_freq),
         any_missing = case_when(is.na(DMDEDUC2) | is.na(URXUCR) | is.na(BMXWAIST) | is.na(LBXCOT) | is.na(DSST) |
                                   is.na(smk_status) |is.na(total_seafood_score) |is.na(alc_month_freq) ~ 1,
                                 .default = 0)) %>%
  select(educ_m, creat_m, waist_m, cot_m, dsst_m, smk_m, seafood_m, alc_m, any_missing)

table(n_miss$any_missing, useNA = 'always')

n_miss %>%
  filter(dsst_m == F) %>%
  count(any_missing)

```

# Included vs excluded
Jul 8: restrict this table to age 60+ only
```{r}

incl_excl_grp <- tbl1_data %>%
  filter(RIDAGEYR >= 60)
  # N = 5466
  
incl_excl_unwt <- tbl_summary(data = incl_excl_grp, 
                            by = 'include', 
                            include = c(DSST, RIDAGEYR, RIAGENDR, RIDRETH1, hs_educ, smk_status, LBXCOT, alc4, 
                                        BMXWAIST, URXUCR, total_seafood_grp, normal_kidney, SDDSRVYR),
                            label = c(DSST ~ 'Digit Symbol Substitution Test (DSST)',
                                      RIDAGEYR ~ 'Age (years)',
                                      RIAGENDR ~ 'Sex',
                                      RIDRETH1 ~ 'Race/ethnicity',
                                      hs_educ ~ 'Education',
                                      smk_status ~ 'Smoking status',
                                      LBXCOT ~ 'Serum cotinine (ng/mL)',
                                      alc4 ~ 'Alcohol consumption',
                                      BMXWAIST ~ 'Waist circumference (cm)',
                                      URXUCR ~ 'Urinary creatinine (mg/dL)',
                                      total_seafood_grp ~ 'Fish and seafood consumption in past 30 days',
                                      normal_kidney ~ 'Estimated glomerular filtration rate (eGFR; ml/min/1.73 m2)',
                                      SDDSRVYR ~ 'NHANES cycle'),
                            missing_text = 'Missing',
                            statistic = list(all_continuous() ~ "{mean} ({sd})",
                                             all_categorical() ~ "{p}%"),
                            digits = list(all_continuous() ~ c(2, 2),
                                          all_categorical() ~ 2)) %>%
  add_overall() %>%
  add_p(list(all_continuous() ~ 't.test',
             all_categorical() ~ 'chisq.test')) %>%
  bold_labels() %>%
  modify_header(label = '**Variable**',
                stat_0 = '**Overall (N = 5,466)**',
                stat_1 = '**Excluded (N = 484)**',
                stat_2 = '**Included (N = 4,982)**') %>%
  modify_caption('Table X. Unweighted descriptive statistics of excluded vs. included participants aged 60 years and older, National Health and Nutrition Examination Survey (NHANES) 1999-2000, 2011-2012, and 2013-2014 (N = 5,466).')

incl_excl_unwt
# need unweighted table to pull the N missing

# weighted
incl_excl_wt_grp <- subset(tbl1_wt_data,
                           RIDAGEYR >= 60)

incl_excl_wt <- tbl_svysummary(data = incl_excl_wt_grp, 
                            by = 'include', 
                            include = c(DSST, RIDAGEYR, RIAGENDR, RIDRETH1, hs_educ, smk_status, LBXCOT, alc4, 
                                        BMXWAIST, URXUCR, total_seafood_grp, normal_kidney, SDDSRVYR),
                            label = c(RIDAGEYR ~ 'Age (years)',
                                      RIAGENDR ~ 'Sex',
                                      RIDRETH1 ~ 'Race/ethnicity',
                                      hs_educ ~ 'Education',
                                      DSST ~ 'Digit Symbol Substitution Test (DSST)',
                                      smk_status ~ 'Smoking status',
                                      LBXCOT ~ 'Serum cotinine (ng/mL)',
                                      alc4 ~ 'Alcohol consumption',
                                      BMXWAIST ~ 'Waist circumference (cm)',
                                      URXUCR ~ 'Urinary creatinine (mg/dL)',
                                      total_seafood_grp ~ 'Fish and seafood consumption in past 30 days',
                                      normal_kidney ~ 'Estimated glomerular filtration rate (eGFR; ml/min/1.73 m2)',
                                      SDDSRVYR ~ 'NHANES cycle'),
                            missing_text = 'Missing',
                            statistic = list(all_continuous() ~ "{mean} ({sd})",
                                             all_categorical() ~ "{p}%"),
                            digits = list(all_continuous() ~ c(2, 2),
                                          all_categorical() ~ 2)) %>%
  add_overall() %>%
  add_p(list(all_continuous() ~ 'svy.t.test',
             all_categorical() ~ 'svy.chisq.test')) %>%
  bold_labels() %>%
  modify_header(label = '**Variable**',
                stat_0 = '**Overall (N = 5,466)**',
                stat_1 = '**Excluded (N = 484)**',
                stat_2 = '**Included (N = 4,982)**') %>%
  modify_caption('Table X. Weighted descriptive statistics of excluded vs. included participants aged 60 years and older, National Health and Nutrition Examination Survey (NHANES) 1999-2000, 2011-2012, and 2013-2014 (N = 5,466).')

incl_excl_wt


incl_excl_unwt %>%
  as_flex_table() %>%
  save_as_docx(path = 'Output Cog ExWAS/Cog ExWAS unweighted descriptive stats incl excl 60plus Jul 8 2024.docx')

incl_excl_wt %>%
  as_flex_table() %>%
  save_as_docx(path = 'Output Cog ExWAS/Cog ExWAS weighted descriptive stats incl excl 60plus Jul 8 2024.docx')

```



