---
title: "Cognition ExWAS run models"
author: "Erika Walker"
date: "`r Sys.Date()`"
output: html_document
---

Purpose: Run cognition ExWAS models in all included participants (cycles 1999-2000, 2011-2012, 2013-2014). Analysis updated to apply NHANES survey weights and multiple imputation

Models: survey-weighted linear regression models for the association between each chemical exposure and DSST

Main model: adjusted for age, sex, race/ethnicity, education, survey cycle, fish consumption, smoking, cotinine, creatinine (if urinary measurement)

Sensitivity analysis 1: restrict to participants with normal eGFR (>=60), main model covariates

Sensitivity analysis 2: add adjustment for waist circumference and alcohol consumption

Sensitivity analysis 3: dichotomize DSST at 25th percentile and run modified Poisson regression

Sensitivity analysis 4: dichotomize DSST at 25th percentile and run modified Poisson regression; restricted to normal eGFR

Jul 9 version: adjust p-values only of the chemical coefficients

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sjlabelled)
library(mitools)
library(mice)
library(survey)
```

# Load and prep data
```{r}
# load imputed dataset
imputed_data <- readRDS('Data/cognition_ewas_svi_imputed.RDS')

# load dictionary of chems to use
use_these_chems <- read_csv('Data/use_these_chems.csv')

chem_list <- use_these_chems$variable_codename_use

# load chemical exposure dataset, filter to cycles 1, 7, 8, select included chems
chems <- read_csv('Data/chemicals_clean.csv', na = '.') %>%
  filter(SDDSRVYR %in% c(1, 7, 8)) %>%
  select(SEQN, SDDSRVYR, all_of(chem_list))
dim(chems) # 27210   149

# correct urinary cadmium = 0 to NA (measurement error in NHANES)
chems <- chems %>%
  mutate(URXUCD = case_when(URXUCD == 0 ~ NA,
                            .default = URXUCD))

# load survey weights, filter to cycles 1, 7, 8, select weights corresponding to included chems
wts_list <- paste('WT_', chem_list, sep = '')

weights <- read_csv('Data/weights_clean.csv', na = '.') %>%
  filter(SDDSRVYR %in% c(1, 7, 8)) %>%
  select(SEQN, SDDSRVYR, all_of(wts_list))
dim(weights) # 29896   149

# set vars as factors
imputed_data <- imputed_data %>%
  as_label(RIAGENDR, hs_educ) %>% # set sex and education as factors
  mutate(RIDRETH1 = relevel(factor(RIDRETH1,
                                   levels = c(1:5),
                                   labels = c('Mexican American',
                                              'Other Hispanic',
                                              'Non-Hispanic White',
                                              'Non-Hispanic Black',
                                              'Other Race')),
                            ref = 'Non-Hispanic White')) %>% # set non hispanic white as reference group
  mutate(cycle = case_when(SDDSRVYR == 1 ~ '1999-2000',
                           SDDSRVYR == 7 ~ '2011-2012',
                           SDDSRVYR == 8 ~ '2013-2014'))

# create variable for mild cognitive impairment: cutoff 25th percentile of DSST
# grab from the analysis group
mci_cutoff<- imputed_data %>%
  filter(include == 1) %>%
  summarize(p25 = quantile(DSST, prob = 0.25)) %>%
  pull(p25)

imputed_data <- imputed_data %>%
  mutate(mild_cog_impair = case_when(include == 1 & DSST < mci_cutoff ~ 1,
                                     include == 1 & DSST >= mci_cutoff ~ 0))

table(imputed_data$mild_cog_impair, useNA = 'always')
# 18987+5923=24910/5=4982, correct sample size
  
table(imputed_data$RIAGENDR)
table(imputed_data$hs_educ)
table(imputed_data$RIDRETH1)
table(imputed_data$SDDSRVYR)
table(imputed_data$cycle)
table(imputed_data$smk_status)
table(imputed_data$include) # indicator for analysis group

# load model functions
source('Cog_ExWAS_model_functions_June_5_2024.R')

```

# Run the models - main analysis
```{r}
cog_models <- chems %>%
  pivot_longer(cols = all_of(chem_list),
               values_to = 'chem_measure',
               names_to = 'chem_name') %>%
  group_by(chem_name) %>%
  do(run_exwas_model(., weights, imputed_data)) %>%
  ungroup()

# clean up results
# add full chem name and family
var_info <- use_these_chems %>%
  select(variable_codename_use, variable_description_use, chemical_family) %>%
  rename(full_chem_name = variable_description_use)

cog_models_clean <- left_join(cog_models,
                              var_info,
                              join_by(chem_name == variable_codename_use))

# reorder columns for easier reading and rename p val
cog_models_clean <- cog_models_clean %>%
  relocate(full_chem_name, .after = chem_name) %>%
  relocate(ci_lower, .after = estimate) %>%
  relocate(ci_upper, .after = ci_lower) %>%
  relocate(p.value, .after = ci_upper) %>%
  rename(p_val_unadj = p.value)

# calculate sample size per chem
chem_n <- chems %>%
  filter(SEQN %in% (imputed_data %>% filter(include == 1))$SEQN) %>%
  pivot_longer(cols = all_of(chem_list),
               values_to = 'chem_measure',
               names_to = 'chem_name') %>%
  drop_na(chem_measure) %>%
  group_by(chem_name) %>%
  summarize(n = n())

cog_models_clean <- left_join(cog_models_clean,
                              chem_n,
                              join_by(chem_name)) %>%
  relocate(n, .after = p_val_unadj)

# false discovery rate correction for chem main effect
cog_models_clean_chem_term <- cog_models_clean %>%
  filter(term == 'scale(log2(chem_measure))') %>%
  mutate(p_val_adj = p.adjust(p_val_unadj, method = 'fdr')) %>%
  relocate(p_val_adj, .after = p_val_unadj)

# merge adj p-values back into overall results
cog_models_clean_adj <- left_join(cog_models_clean,
                                  cog_models_clean_chem_term,
                                  by = intersect(colnames(cog_models_clean), 
                                                 colnames(cog_models_clean_chem_term))) %>%
  relocate(p_val_adj, .after = p_val_unadj)


```

## save output
```{r}

write_csv(cog_models_clean_adj, file = 'Output Cog ExWAS/Cog ExWAS main analysis model output Jul 9 2024.csv')
write_csv(cog_models_clean_chem_term, file = 'Output Cog ExWAS/Cog ExWAS main analysis model output chem term Jul 9 2024.csv')

```

# Sensitivity analysis: normal eGFR subset
```{r}

summary((imputed_data %>% filter(include == 1))$VNEGFR)

# create indicator for eGFR >= 60
imputed_data <- imputed_data %>%
  mutate(egfr_grp = case_when(include == 1 & VNEGFR >= 60 ~ 1,
                              .default = 0))

table(imputed_data$egfr_grp, useNA = 'always')
# 17965/5=3593

cog_models_egfr <- chems %>%
  pivot_longer(cols = all_of(chem_list),
               values_to = 'chem_measure',
               names_to = 'chem_name') %>%
  group_by(chem_name) %>%
  do(run_exwas_model_normal_egfr(., weights, imputed_data)) %>%
  ungroup()

# clean up results
# add full chem name and family
cog_models_egfr_clean <- left_join(cog_models_egfr,
                                   var_info,
                                   join_by(chem_name == variable_codename_use))

# reorder columns for easier reading and rename p val
cog_models_egfr_clean <- cog_models_egfr_clean %>%
  relocate(full_chem_name, .after = chem_name) %>%
  relocate(ci_lower, .after = estimate) %>%
  relocate(ci_upper, .after = ci_lower) %>%
  relocate(p.value, .after = ci_upper) %>%
  rename(p_val_unadj = p.value)

# add sample size per chem
chem_n_egfr <- chems %>%
  filter(SEQN %in% (imputed_data %>% filter(egfr_grp == 1))$SEQN) %>%
  pivot_longer(cols = all_of(chem_list),
               values_to = 'chem_measure',
               names_to = 'chem_name') %>%
  drop_na(chem_measure) %>%
  group_by(chem_name) %>%
  summarize(n = n())

cog_models_egfr_clean <- left_join(cog_models_egfr_clean,
                                   chem_n_egfr,
                                   join_by(chem_name)) %>%
  relocate(n, .after = p_val_unadj)

# false discovery rate correction for chem main effect
cog_models_egfr_clean_chem_term <- cog_models_egfr_clean %>%
  filter(term == 'scale(log2(chem_measure))') %>%
  mutate(p_val_adj = p.adjust(p_val_unadj, method = 'fdr')) %>%
  relocate(p_val_adj, .after = p_val_unadj)

# merge adj p-values back into overall results
cog_models_egfr_clean_adj <- left_join(cog_models_egfr_clean,
                                       cog_models_egfr_clean_chem_term,
                                       by = intersect(colnames(cog_models_egfr_clean), 
                                                      colnames(cog_models_egfr_clean_chem_term))) %>%
  relocate(p_val_adj, .after = p_val_unadj)

```

## save output
```{r}

write_csv(cog_models_egfr_clean_adj, file = 'Output Cog ExWAS/Cog ExWAS normal eGFR model output Jul 9 2024.csv')
write_csv(cog_models_egfr_clean_chem_term, file = 'Output Cog ExWAS/Cog ExWAS normal eGFR model output chem term Jul 9 2024.csv')

```

# Sensitivity analysis: add waist circumference and alcohol consumption
```{r}

cog_models_lifestyle <- chems %>%
  pivot_longer(cols = all_of(chem_list),
               values_to = 'chem_measure',
               names_to = 'chem_name') %>%
  group_by(chem_name) %>%
  do(run_exwas_model_lifestyle(., weights, imputed_data)) %>%
  ungroup()

# clean up results
# add full chem name and family
cog_models_lifestyle_clean <- left_join(cog_models_lifestyle,
                                        var_info,
                                        join_by(chem_name == variable_codename_use))

# reorder columns for easier reading and rename p val
cog_models_lifestyle_clean <- cog_models_lifestyle_clean %>%
  relocate(full_chem_name, .after = chem_name) %>%
  relocate(ci_lower, .after = estimate) %>%
  relocate(ci_upper, .after = ci_lower) %>%
  relocate(p.value, .after = ci_upper) %>%
  rename(p_val_unadj = p.value)

# add sample size per chem - will be same as main analysis
cog_models_lifestyle_clean <- left_join(cog_models_lifestyle_clean,
                                        chem_n,
                                        join_by(chem_name)) %>%
  relocate(n, .after = p_val_unadj)

# false discovery rate correction for chem main effect
cog_models_lifestyle_clean_chem_term <- cog_models_lifestyle_clean %>%
  filter(term == 'scale(log2(chem_measure))') %>%
  mutate(p_val_adj = p.adjust(p_val_unadj, method = 'fdr')) %>%
  relocate(p_val_adj, .after = p_val_unadj)

# merge adj p-values back into overall results
cog_models_lifestyle_clean_adj <- left_join(cog_models_lifestyle_clean,
                                       cog_models_lifestyle_clean_chem_term,
                                       by = intersect(colnames(cog_models_lifestyle_clean),
                                                      colnames(cog_models_lifestyle_clean_chem_term))) %>%
  relocate(p_val_adj, .after = p_val_unadj)

```

## save output
```{r}

write_csv(cog_models_lifestyle_clean_adj, file = 'Output Cog ExWAS/Cog ExWAS lifestyle model output Jul 9 2024.csv')
write_csv(cog_models_lifestyle_clean_chem_term, file = 'Output Cog ExWAS/Cog ExWAS lifestyle model output chem term Jul 9 2024.csv')

```

# Sensitivity analysis: binary cognition outcome
```{r}

cog_models_poisson <- chems %>%
  pivot_longer(cols = all_of(chem_list),
               values_to = 'chem_measure',
               names_to = 'chem_name') %>%
  group_by(chem_name) %>%
  do(run_exwas_model_poisson(., weights, imputed_data)) %>%
  ungroup()

# clean up results
# add full chem name and family
cog_models_poisson_clean <- left_join(cog_models_poisson,
                                      var_info,
                                      join_by(chem_name == variable_codename_use))

# reorder columns for easier reading and rename p val
cog_models_poisson_clean <- cog_models_poisson_clean %>%
  relocate(full_chem_name, .after = chem_name) %>%
  relocate(ci_lower, .after = estimate) %>%
  relocate(ci_upper, .after = ci_lower) %>%
  relocate(p.value, .after = ci_upper) %>%
  rename(p_val_unadj = p.value)

# add sample size per chem - will be same as main analysis
cog_models_poisson_clean <- left_join(cog_models_poisson_clean,
                                      chem_n,
                                      join_by(chem_name)) %>%
  relocate(n, .after = p_val_unadj)

# false discovery rate correction for chem main effect
cog_models_poisson_clean_chem_term <- cog_models_poisson_clean %>%
  filter(term == 'scale(log2(chem_measure))') %>%
  mutate(p_val_adj = p.adjust(p_val_unadj, method = 'fdr')) %>%
  relocate(p_val_adj, .after = p_val_unadj)

# merge adj p-values back into overall results
cog_models_poisson_clean_adj <- left_join(cog_models_poisson_clean,
                                       cog_models_poisson_clean_chem_term,
                                       by = intersect(colnames(cog_models_poisson_clean),
                                                      colnames(cog_models_poisson_clean_chem_term))) %>%
  relocate(p_val_adj, .after = p_val_unadj)

```

## save output
```{r}

write_csv(cog_models_poisson_clean_adj, file = 'Output Cog ExWAS/Cog ExWAS poisson model output Jul 9 2024.csv')
write_csv(cog_models_poisson_clean_chem_term, file = 'Output Cog ExWAS/Cog ExWAS poisson model output chem term Jul 9 2024.csv')

```

# Sensitivity analysis: binary cognition outcome, normal eGFR
```{r}

cog_models_poisson_egfr <- chems %>%
  pivot_longer(cols = all_of(chem_list),
               values_to = 'chem_measure',
               names_to = 'chem_name') %>%
  group_by(chem_name) %>%
  do(run_exwas_model_poisson_egfr(., weights, imputed_data)) %>%
  ungroup()

# clean up results
# add full chem name and family
cog_models_poisson_egfr_clean <- left_join(cog_models_poisson_egfr,
                                      var_info,
                                      join_by(chem_name == variable_codename_use))

# reorder columns for easier reading and rename p val
cog_models_poisson_egfr_clean <- cog_models_poisson_egfr_clean %>%
  relocate(full_chem_name, .after = chem_name) %>%
  relocate(ci_lower, .after = estimate) %>%
  relocate(ci_upper, .after = ci_lower) %>%
  relocate(p.value, .after = ci_upper) %>%
  rename(p_val_unadj = p.value)

# add sample size per chem - will be same as main egfr analysis
cog_models_poisson_egfr_clean <- left_join(cog_models_poisson_egfr_clean,
                                      chem_n_egfr,
                                      join_by(chem_name)) %>%
  relocate(n, .after = p_val_unadj)

# false discovery rate correction for chem main effect
cog_models_poisson_egfr_clean_chem_term <- cog_models_poisson_egfr_clean %>%
  filter(term == 'scale(log2(chem_measure))') %>%
  mutate(p_val_adj = p.adjust(p_val_unadj, method = 'fdr')) %>%
  relocate(p_val_adj, .after = p_val_unadj)

# merge adj p-values back into overall results
cog_models_poisson_egfr_clean_adj <- left_join(cog_models_poisson_egfr_clean,
                                       cog_models_poisson_egfr_clean_chem_term,
                                       by = intersect(colnames(cog_models_poisson_egfr_clean),
                                                      colnames(cog_models_poisson_egfr_clean_chem_term))) %>%
  relocate(p_val_adj, .after = p_val_unadj)

```

## save output
```{r}

write_csv(cog_models_poisson_egfr_clean_adj, file = 'Output Cog ExWAS/Cog ExWAS poisson normal eGFR model output Jul 9 2024.csv')
write_csv(cog_models_poisson_egfr_clean_chem_term, file = 'Output Cog ExWAS/Cog ExWAS poisson normal eGFR model output chem term Jul 9 2024.csv')

```
